---
title: "airMONTE"
output: pdf_document
---

```{r  reproducibility}
set.seed(400)
```

```{r libraries, message= FALSE}
library(dplyr)
library(stats)
library(tidyr)
library(tidyverse)
library(broom)
library(ggplot2)
```


```{r setting}
m <- 10000

monte_AIR <- function(m) {
  storing <- numeric(m)
  for (i in 1:m){
    p <- rnorm(1, mean = 0.48, sd = 0.20)
    I <- 1
    phi <- rnorm(1, mean = 100, sd = 20)
    V <- 132
    N <- rnorm(1, mean = 3, sd = 0.5)
    t <- rnorm(1, mean = 0.5, sd = 0.167)
  
    form <- ((p*I*phi)/V) * ((N*t + exp(-N*t) - 1) / (N^2))
  
    storing[i] <- 1 - exp(-form)
  }
  return(storing)
}

we_want <- monte_AIR(m)

mean(we_want)

hist(we_want , breaks = 35, main = "Monte Carlo Air Replication")
```


```{r}
# Wells-Riley formula: 

## vent_mean = 4 <- ventilation rate
## breath_mean = 0.48 <- pumonary ventilation rate
## time_mean = 0.5 <- 30min /1 hr

air_monte <- function(m = 10000, quanta_mean, quanta_sd, 
                      room_vol=132, infect=1, vent_mean=4, vent_sd=1, 
                      breath_mean=0.48, breath_sd=0.2, 
                      time_mean=0.5, time_sd=0.167, s_mean = 19, s_sd=10){
  prob_vec <- numeric(m)
  infect_vect <- numeric(m)
  s_after <- numeric(m)
  
  for (i in 1:m) {
    p <- max(rnorm(1, breath_mean, breath_sd), 0)
    phi <- max(rnorm(1, quanta_mean, quanta_sd), 0)
    t <- max(rnorm(1, time_mean, time_sd), 0)
    Q <- max(rnorm(1, vent_mean, venth_sd), 0.01)
    S0 <- max(round(rnorm(1, s_mean, s_sd)), 1)
    
      # exposure term
    X <- (infect * phi * p * t) / Q
    
      # infection prob. 
    P <- 1 - exp(-X)
    
      # expected infections
    exp_inf <- S0 * P
    
    prob_vec[i] <- P
    infect_vect[i] <- exp_inf
    s_after[i] <- S0 - exp_inf
  }
 
  return(list(
    prob=prob_vec, infections=infect_vect, non_infected = s_after,
    summary = list(
      mean_prob = mean(prob_vec),
      mean_infections = mean(inf_vect),
      prob_quantiles = quantile(prob_vec, c(0.25,0.5,0.75,0.9,0.95)),
      inf_quantiles = quantile(infect_vect, c(0.25,0.5,0.75,0.9,0.95))
    )
  ))
}




```







## For the different quantas:
TB: mean = 12.7,  SD = 3.0
Flu: mean = 100,  SD = 25.0
Measles: mean = 570,  SD = 143.0

Waiting room Volume: 132 (m^3)

Mean Pulmonary Ventilation Rate: mean = 0.48, SD = 0.2
Mean Waiting Time: t = 30 min,  SD = 10 min
Room Ventilation Rate: mean = 4
Susceptibles Present: mean = 19,  SD = 10

but I = 1 infectious individual per simulation


```{r Tuberculosis}
better_monte <- function(m = 10000, dis_quanta, quanta_sd, room_size = 132, 
                         infected_individuals = 1, room_venti_rate = 4){ 
  probability_inf <- numeric(m)
  num_new_infect <- numeric(m)
  sus_persons <- numeric(m)

  for(i in 1:m) {
    
    p_2 <- max(rnorm(1, mean = 0.48, sd = 0.20), 0)
    phi_2 <- max(rnorm(1, mean = dis_quanta, sd = quanta_sd),0)
    t_2 <- max(rnorm(1, mean = 0.5, sd = 0.167), 0)
    S0 <- max(round(rnorm(1, mean = 19, sd = 10)),1)
    
    q_2 <- infected_individuals * phi_2
 
    n_t <- (q_2/room_venti_rate) * (1 - exp(-room_venti_rate*t_2))
    
    nucci <- ((p_2*q_2)/room_size) * 
      ((room_venti_rate * t_2 + exp(-room_venti_rate * 
                                      t_2) -1) / (room_venti_rate^2))
    
    S_t <- S0 * exp(-nucci)
    
    inf <- S0 - S_t
    
    sus_prob <- inf / S0
    
    probability_inf[i] <- sus_prob
    num_new_infect[i] <- inf
    sus_persons[i] <- S_t
  }
  
  return(list(prob = probability_inf,
              infections = num_new_infect,
              survivors = sus_persons,
              summary = list(mean_prob = mean(probability_inf),
                             mean_infections = mean(num_new_infect),
                             prob_quantiles = quantile(probability_inf, 
                                                       c(0.25, 0.5, 0.75, 
                                                         0.9, 0.95)),
                             inf_quantiles = quantile(num_new_infect, 
                                                      c(0.25, 0.5, 0.75, 
                                                        0.9, 0.95)))))

}

TB <- better_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_size = 132, infected_individuals = 1, 
                   room_venti_rate = 4)

mean(TB$prob)
h <- hist(TB$prob, breaks = 35, main = "TB Replication")

h$counts <- h$counts / sum(h$counts) * 100

plot(h, main = "TBL Prob Freq Distrib")


dens <- density(TB$prob)

dens$y <- dens$y / sum(dens$y) *100

plot(dens, main = "TB: Prob Freq Distrib 2")
```


```{r Measles}
Measles <- better_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_size = 132, infected_individuals = 1, 
                   room_venti_rate = 4)

mean(Measles$prob)
hMeas <- hist(Measles$prob, breaks = 35, main = "Measles Replication")

hMeas$counts <- hMeas$counts / sum(hMeas$counts) * 100

plot(h, main = "Measles Prob Freq Distrib")


densM <- density(Measles$prob)

densM$y <- densM$y / sum(densM$y) *100

plot(densM, main = "Measles: Prob Freq Distrib 2")
```

```{r Influenza}
Influenza <- better_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_size = 132, infected_individuals = 1, 
                   room_venti_rate = 4)

mean(Influenza$prob)

hFLU <- hist(Influenza$prob, breaks = 35, main = "Influenza Replication")

hFLU$counts <- hFLU$counts / sum(hFLU$counts) * 100

plot(hFLU, main = "Influenza Prob Freq Distrib")


densFLU <- density(Influenza$prob)

densFLU$y <- densFLU$y / sum(densFLU$y) *100

plot(densFLU, main = "Influenza: Prob Freq Distrib 2")

```








In their study: occupancy levels, waiting times, and ventilation rate varied.



*Wanting to compare COVID/SARS quantum to the already listed diseases
Quantum: generation rate of 18-48 h^-1  for COVID-19
Infection prob: less than 1% requires ventilation rate > 100-350 m^3/h per infector, and 1200-4000 m^3/h per infector for 0.25 h and 3 h of exposure
Wells-Riley equation: $P = C/S = 1-e^{Iqpt/Q}$

https://link.springer.com/article/10.1007/s12273-020-0703-5 

```{r second disease: SARS}
sars_mean <- 2.7  #mean quanta per hour
sars_sd <- 1.2

sars <- better_monte(m, dis_quanta = sars_mean, quanta_sd = sars_sd,
                   room_size = 132, infected_individuals = 1, room_venti_rate = 4)
summary(sars)
mean(sars$prob)

```


