---
title: "airMONTE"
output: pdf_document
---

```{r  reproducibility}
set.seed(400)
```

```{r libraries, message= FALSE}
library(dplyr)
library(stats)
library(tidyr)
library(tidyverse)
library(broom)
library(ggplot2)
library(tibble)
```


```{r setting}
m <- 10000

monte_AIR <- function(m) {
  storing <- numeric(m)
  for (i in 1:m){
    p <- rnorm(1, mean = 0.48, sd = 0.20)
    I <- 1
    phi <- rnorm(1, mean = 100, sd = 20)
    V <- 132
    N <- rnorm(1, mean = 3, sd = 0.5)
    t <- rnorm(1, mean = 0.5, sd = 0.167)
  
    form <- ((p*I*phi)/V) * ((N*t + exp(-N*t) - 1) / (N^2))
  
    storing[i] <- 1 - exp(-form)
  }
  return(storing)
}

we_want <- monte_AIR(m)

mean(we_want)

hist(we_want , breaks = 35, main = "Monte Carlo Air Replication")
```


```{r}
# Wells-Riley formula: 

## vent_mean = 4 <- ventilation rate
## breath_mean = 0.48 <- pumonary ventilation rate
## time_mean = 0.5 <- 30min /1 hr

air_monte <- function(m = 10000, quanta_mean, quanta_sd, 
                      room_vol=132, infect=1, vent_mean=4, vent_sd=1, 
                      breath_mean=0.48, breath_sd=0.2, 
                      time_mean=0.5, time_sd=0.167, s_mean = 19, s_sd=10) {

  
    p_vec <- pmax(rnorm(m, breath_mean, breath_sd), 0)
    phi_vec <- pmax(rnorm(m, quanta_mean, quanta_sd), 0)
    t_vec <- pmax(rnorm(m, time_mean, time_sd), 0)
    Q_vec <- pmax(rnorm(m, vent_mean, vent_sd), 0.01)
    S0_vec <- pmax(round(rnorm(m, s_mean, s_sd)), 1)
    I_vec <- infect
    
      # exposure term
    X_vec <- (I_vec * phi_vec * p_vec * t_vec) / Q_vec
    
      # infection prob. 
    P_vec <- 1 - exp(-X_vec)
    
    new_infect <- S0_vec * P_vec
    S_after <- S0_vec - new_infect
    
    res_df <- tibble::tibble(
      p = p_vec, phi=phi_vec, t=t_vec, Q=Q_vec, S0=S0_vec, 
      X=X_vec, prob=P_vec, infections=new_infect, survivors=S_after
    )
    
    summ <- list(
      mean_prob = mean(res_df$prob),
      mean_infections = mean(res_df$infections),
      prob_quantiles = quantile(res_df$prob, c(0.25,0.5,0.75,0.9,0.95)),
      inf_quantiles = quantile(res_df$infections, c(0.25,0.5,0.75,0.9,0.95))
    )
    return(list(data=res_df, summary=summ))
}

plot_prob <- function(sim_res, title="Infection Prob. Distribution"){
  df <- sim_res$data
  ggplot(df, aes(x=prob)) +
    geom_histogram(aes(y=..count.. / sum(..count..) * 100), bins=35, closed="left") +
    labs(x="Infection Prob. for an Individual(P)", 
         y="Percent of Simulation (%)",
         title=title)
}
```

```{r testing model}
TB <- air_monte(m=10000, quanta_mean=12.7, quanta_sd=3.0)
Flu <- air_monte(m=10000, quanta_mean=100, quanta_sd=25.0)
Meas <- air_monte(m=10000, quanta_mean=570, quanta_sd=143.0)
SARS <- air_monte(m=10000, quanta_mean=2.7, quanta_sd=1.2)

TB

```




## For the different quantas:
TB: mean = 12.7,  SD = 3.0
Flu: mean = 100,  SD = 25.0
Measles: mean = 570,  SD = 143.0

Waiting room Volume: 132 (m^3)

Mean Pulmonary Ventilation Rate: mean = 0.48, SD = 0.2
Mean Waiting Time: t = 30 min,  SD = 10 min
Room Ventilation Rate: mean = 4
Susceptible Patients: mean = 19,  SD = 10

but I = 1 infectious individual per simulation


```{r Tuberculosis}
better_monte <- function(m = 10000, dis_quanta, quanta_sd, room_vol = 132, 
                         infected_individuals = 1, ACH = 4){ 
  probability_inf <- numeric(m)
  num_new_infect <- numeric(m)
  sus_persons <- numeric(m)

  for(i in 1:m) {
    
    p_2 <- max(rnorm(1, mean = 0.48, sd = 0.20), 0)
    phi_2 <- max(rnorm(1, mean = dis_quanta, sd = quanta_sd),0)
    t_2 <- max(rnorm(1, mean = 0.5, sd = 0.167), 0)
    S0 <- max(round(rnorm(1, mean = 19, sd = 10)),1)
    
    q_2 <- infected_individuals * phi_2
 
    
    nucci <- ((p_2*q_2)/room_vol) * 
      ((ACH * t_2 + exp(-ACH *t_2) -1) / (ACH^2))
    
    S_t <- S0 * exp(-nucci)
    
    
    exp_inf <- S0 - S_t
    inf <- rpois(1, lambda = exp_inf)
    
    sus_prob <- inf / S0
    
    probability_inf[i] <- sus_prob
    num_new_infect[i] <- inf
    sus_persons[i] <- S_t
  }
  
  return(list(prob = probability_inf,
              infections = num_new_infect,
              survivors = sus_persons,
              summary = list(mean_prob = mean(probability_inf),
                             mean_infections = mean(num_new_infect),
                             prob_quantiles = quantile(probability_inf, 
                                                       c(0.25, 0.5, 0.75, 
                                                         0.9, 0.95)),
                             inf_quantiles = quantile(num_new_infect, 
                                                      c(0.25, 0.5, 0.75, 
                                                        0.9, 0.95)))))

}

TB <- better_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

TB

mean(TB$prob)

TB_df <- data.frame(prob = TB$prob)

TB_plot <- ggplot(data = TB_df, aes(x = prob)) + 
  geom_histogram(aes(y = ..count.. /sum(..count..) *100),
                 bins = 35, 
                 fill = "lightgrey",
                 color = "black") +
  labs(title = "TB Prob. Distrib.",
       x = "Prob. of Infection",
       y = "Percentage of Simulations")

TB_plot

```

```{r Flu}
Flu <- better_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(Flu$prob)
hMeas <- hist(Flu$prob, breaks = 35, main = "Measles Replication")

hFLu$counts <- hFlu$counts / sum(hFlu$counts) * 100

plot(h, main = "Flu Prob Freq Distrib")


densF <- density(Flu$prob)

densF$y <- densF$y / sum(densF$y) *100

plot(densF, main = "Flu: Prob Freq Distrib 2")
```

```{r Measles}
Meas <- better_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(Meas$prob)

hM <- hist(Meas$prob, breaks = 35, main = "Measles Replication")

hM$counts <- hM$counts / sum(hM$counts) * 100

plot(hM, main = "Measles Prob Freq Distrib")


densM <- density(Meas$prob)

densM$y <- densM$y / sum(densM$y) *100

plot(densM, main = "Measles: Prob Freq Distrib 2")

```


```{r SARS}
SARS <- better_monte(10000, dis_quanta = 2.7, quanta_sd = 1.2,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(SARS$prob)

hSARS <- hist(SARS$prob, breaks = 35, main = "SARS Graph")

hSARS$counts <- hSARS$counts / sum(hSARS$counts) * 100

plot(hSARS, main = "SARS Prob Freq Distrib")


densSARS <- density(SARS$prob)

densSARS$y <- densSARS$y / sum(densSARS$y) *100

plot(densSARS, main = "SARS: Prob Freq Distrib 2")

```


## Extended Section:
(Where we run through SARS quanta and compare it to the 3 other diseases from 
the study)
Reasoning: To see the results of a different airborne disease and compare it to 
the other disease results.

and 

(Where we're running different simulations, messing with exposure time, # of
infected individuals, and room size)
Reasoning: Since waiting rooms are not a universal size, exposure times are not 
set, and # of infected individuals can vary, it makes sense to investigate 
the results that come from these changes.

Scenario 1:
Room size is bigger than the one from the study
Exposure time is longer?
The number of infected individuals is the same

Prediction: smaller probability


Scenario 2:
Room size is same than the one from the study
Exposure time is longer
The number of infected individuals is 2?

Prediction: larger probability

Scenario 3:
Room size is bigger.
Exposure time is shorter
# of infected individuals is larger. 

Prediction: smaller probability






In their study: occupancy levels, waiting times, and ventilation rate varied.

*Wanting to compare COVID/SARS quantum to the already listed diseases
Quantum: generation rate of 18-48 h^-1  for COVID-19
Infection prob: less than 1% requires ventilation rate > 100-350 m^3/h per infector, and 1200-4000 m^3/h per infector for 0.25 h and 3 h of exposure
Wells-Riley equation: $P = C/S = 1-e^{Iqpt/Q}$

https://link.springer.com/article/10.1007/s12273-020-0703-5 



```{r summary results}
sum_disease <- tibble::tibble(
  Disease=c("TB","Flu", "Meas", "SARS"),
  Mean_Prob = c(TB$summary$mean_prob,
                Flu$summary$mean_prob,
                Meas$summary$mean_prob,
                SARS$summary$mean_prob),
  Mean_Infections = c(TB$summary$mean_infections,
                      Flu$summary$mean_infections,
                      Meas$summary$mean_infections,
                      SARS$summary$mean_infections),
  Prob_25 = c(TB$summary$prob_quantiles[1],
              Flu$summary$prob_quantiles[1],
              Meas$summary$prob_quantiles[1],
              SARS$summary$prob_quantiles[1]),
  Prob_50 = c(TB$summary$prob_quantiles[2],
              Flu$summary$prob_quantiles[2],
              Meas$summary$prob_quantiles[2],
              SARS$summary$prob_quantiles[2]),
  Prob_75 = c(TB$summary$prob_quantiles[3],
              Flu$summary$prob_quantiles[3],
              Meas$summary$prob_quantiles[3],
              SARS$summary$prob_quantiles[3]),
  Prob_90 = c(TB$summary$prob_quantiles[4],
              Flu$summary$prob_quantiles[4],
              Meas$summary$prob_quantiles[4],
              SARS$summary$prob_quantiles[4]),
  Prob_95 = c(TB$summary$prob_quantiles[5],
              Flu$summary$prob_quantiles[5],
              Meas$summary$prob_quantiles[5],
              SARS$summary$prob_quantiles[5])
)

sum_dis_round <- sum_disease %>%
  mutate(across(where(is.numeric), ~ round(.x,3)))

sum_dis_round

```

```{r summary}

## The monte carlo simulation results estimate the infection risk for four airborne disease in a standard waiting room scenario. 

## Tuberculosis: moderate mean infection prob of 0.499 per susceptible individual, with approx 9.6 expected new infections, high variability: range from 0.342 to 0.857, showing potential higher-risk scenarios in dif conditions

## Influenza: higher mean prob: 0.941, nearly 18 expected infections, median and upper quantiles approach 1, infection is almost certain in the environment

## Measles: highest risk, mean prob. of 0.987, almost entire susceptible pop expected to become infected, all quantiles 1

## SARS: low mean prob: 0.153, roughly 3 expected new infections, moderate variability ranges from 0.072 to 0.369

## *infection risk is strongly driven by the disease'airborne quanta generation rate: meas and flu-high risk, tb-moderate risk, sars-lower risk

```



```{r graph}

TB_df <- data.frame(prob=as.numeric(TB$prob), Disease="TB")
Measles_df <- data.frame(prob=as.numeric(Measles$prob), Disease = "Measles")
Influenza_df <- data.frame(prob=as.numeric(Influenza$prob), Disease = "Influenza")
SARS_df <- data.frame(prob=as.numeric(SARS$prob), Disease="SARS")

combined_df <- rbind(TB_df, Measles_df, Influenza_df, SARS_df)
combined_df$prob <- pmin(pmax(combined_df$prob, 0),1)

ggplot(combined_df, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y= ..count.. / sum(..count..) * 100),
                  bins=35, position="identity", alpha=0.4) +
  labs(title="Prob. of Infection Distribution by Disease",
       x = "Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue", "SARS"="yellow3")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue", "SARS"="yellow3")) +
  theme_minimal()

```


```{r Scenario 1 TB}
scenario_monte <- function(m = 10000, dis_quantaS, quanta_sdS, room_vol = 300, 
                         infected_individuals = 1, ACH = 4, t_scen = max(rnorm(1, mean = 1, sd = 0.167))){ 
  prob_inf <- numeric(m)
  num_infec <- numeric(m)
  sus_per <- numeric(m)

  for(i in 1:m) {
    
    p_scen <- max(rnorm(1, mean = 0.48, sd = 0.20), 0)
    phi_scen <- max(rnorm(1, mean = dis_quantaS, sd = quanta_sdS),0)
    S0_scen <- max(round(rnorm(1, mean = 19, sd = 10)),1)
    
    q_2_scen <- infected_individuals * phi_scen
 
    nucci_scen <- ((p_scen*q_2_scen)/room_vol) * 
      ((ACH * t_scen + exp(-ACH * t_scen) -1) / (ACH^2))
    
    S_t_scen <- S0_scen * exp(-nucci_scen)
    
    exp_inf_scen <- S0_scen - S_t_scen
    inf_scen <- rpois(1, lambda = exp_inf_scen)
    
    sus_prob_scen <- inf_scen / S0_scen
    
    prob_inf[i] <- sus_prob_scen
    num_infec[i] <- inf_scen
    sus_per[i] <- S_t_scen
  }
  
  return(list(prob = prob_inf,
              infections = num_infec,
              survivors = sus_per,
              summary = list(mean_prob = mean(prob_inf),
                             mean_infections = mean(num_infec),
                             prob_quantiles = quantile(prob_inf, 
                                                       c(0.25, 0.5, 0.75, 
                                                         0.9, 0.95)),
                             inf_quantiles = quantile(num_infec, 
                                                      c(0.25, 0.5, 0.75, 
                                                        0.9, 0.95)))))

}

scenario_1 <- scenario_monte(10000, dis_quantaS = 12.7, quanta_sdS = 3.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4, t_scen = max(rnorm(1, mean = 1, sd = 0.167)))

mean(scenario_1$prob)
```

```{r Scenario 1 Graph}
scen1_TB <- scenario_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)
scen1_meas <- scenario_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)
scen1_flu <- scenario_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)

TB_df1 <- data.frame(prob=as.numeric(scen1_TB$prob), Disease="TB")
Meas_df1 <- data.frame(prob=as.numeric(scen1_meas$prob), Disease="Measles")
Flu_df1 <- data.frame(prob=as.numeric(scen1_flu$prob), Disease="Influenza")

comb_df1 <- rbind(TB_df1, Meas_df1, Flu_df1)
comb_df1$prob <- pmin(pmax(comb_df1$prob, 0),1)

ggplot(comb_df1, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y=..count.. / sum(..count..) * 100),
                 bins=35, position="identity", alpha=0.4) +
  labs(title="Scenario 1: Probability of Infection Distribution by Disease",
       x="Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  theme_minimal()


```





Scenario 2:
Room volume is 132
Exposure time is 30 min
The number of infected individuals is 2

Prediction: larger probability


```{r Scenario 2 TB}

scenario_2 <- scenario_monte(10000, dis_quantaS = 12.7, quanta_sdS = 3.0,
                             room_vol = 132, infected_individuals = 2, ACH = 4,
                             t_scen = max(rnorm(1, mean = 0.5, sd = 0.167)))

mean(scenario_2$prob)

```

```{r Scenario 1 Measles}
scen_1_M <- scenario_monte(10000, dis_quantaS = 570, quanta_sdS = 143.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4, t_scen = max(rnorm(1, mean = 1, sd = 0.167)))

mean(scen_1_M$prob)
```

```{r Scenario 2 Measles}

scen_2_M <- scenario_monte(10000, 570, 143.0, 132, 2, 4, max(rnorm(1, mean = 0.5, sd = 0.167)))

mean(scen_2_M$prob)
```

## For the different quantas:
TB: mean = 12.7,  SD = 3.0
Flu: mean = 100,  SD = 25.0
Measles: mean = 570,  SD = 143.0

```{r Scenario 1 Flu}
s1_Flu <- scenario_monte(10000, 100, 25.0, 300, 1, 4, 
                         max(rnorm(1, mean = 1, sd = 0.167)))

mean(s1_Flu$prob)
```

```{r Scenario 2 Flu}

s2_Flu <- scenario_monte(10000, 100, 25.0, 132, 2, 4, 
                         max(rnorm(1, mean = 0.5, sd = 0.167)))
mean(s2_Flu$prob)
```


```{r Scenario 2 Graph}
scen2_TB <- scenario_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)
scen2_meas <- scenario_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)
scen2_flu <- scenario_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)

TB_df2 <- data.frame(prob=as.numeric(scen2_TB$prob), Disease="TB")
Meas_df2 <- data.frame(prob=as.numeric(scen2_meas$prob), Disease="Measles")
Flu_df2 <- data.frame(prob=as.numeric(scen2_flu$prob), Disease="Influenza")

comb_df2 <- rbind(TB_df2, Meas_df2, Flu_df2)
comb_df2$prob <- pmin(pmax(comb_df2$prob, 0),1)

ggplot(comb_df2, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y=..count.. / sum(..count..) * 100),
                 bins=35, position="identity", alpha=0.4) +
  labs(title="Scenario 2: Probability of Infection Distribution by Disease",
       x="Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  theme_minimal()

```



