---
title: "Monte Carlo Applied to Infectious Airborne Diseases"
subtitle: "Stat 400"
author: "Group 1: Robin Thrush and Kaitlynn Walston"
output: 
  ioslides_presentation:
    css: styles.css
    indentation: TRUE
---

```{r setup, include=FALSE}
set.seed(445)
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(stats)
library(tidyr)
library(tidyverse)
library(broom)
library(ggplot2)
library(tibble)
```
## Introduction

Stochastic Monte Carlo Model to simulate the probability of a susceptible person
catching an airborne disease.


## Article

Beggs. et al., a study conducted to understand the harm infectious people bring to healthcare waiting rooms. Their
hope was to to have findings that backed up the idea to have more precautions to minimize the spread of airborne infections. The diseases they looked at were Tuberculosis, Measles, and Influenza.

Assumptions they made:
- Only one infectious individual present in each simulation
- Waiting room is ventilated w outside air at a constant flow rate 
- Air in waiting room is mixed so infectious particles are evenly distributed throughout the room space.
- values of p, phi, t, and S vary and are normally distributed.
- applied an SD of 25% of the mean quanta production val. 
- Once susceptible individuals become infected they cannot infect anyone else (bc they leave within the 30min they are there.)


## Equation:
\[
P = 1 - \exp \Bigg[ - \frac{p \, I \, \phi}{V} \cdot
\frac{N t + e^(-N t) - 1}{N^2} \Bigg]
\]


## Replication

```{r Replication, echo=FALSE}
better_monte <- function(m = 10000, dis_quanta, quanta_sd, room_vol = 132, 
                         infected_individuals = 1, ACH = 4){ 
  probability_inf <- numeric(m)
  num_new_infect <- numeric(m)
  sus_persons <- numeric(m)

  for(i in 1:m) {
    
    p_2 <- max(rnorm(1, mean = 0.48, sd = 0.20), 0)
    phi_2 <- max(rnorm(1, mean = dis_quanta, sd = quanta_sd),0)
    t_2 <- max(rnorm(1, mean = 0.5, sd = 0.167), 0)
    S0 <- max(round(rnorm(1, mean = 19, sd = 10)),1)
    
    q_2 <- infected_individuals * phi_2
 
    nucci <- ((p_2*q_2)/room_vol) * 
      ((ACH * t_2 + exp(-ACH * 
                                      t_2) -1) / (ACH^2))
    
    S_t <- S0 * exp(-nucci)
    
    exp_inf <- S0 - S_t
    inf <- rpois(1, lambda = exp_inf)
    
    sus_prob <- inf / S0
    
    probability_inf[i] <- sus_prob
    num_new_infect[i] <- inf
    sus_persons[i] <- S_t
  }
  
  return(list(prob = probability_inf,
              infections = num_new_infect,
              survivors = sus_persons,
              summary = list(mean_prob = mean(probability_inf),
                             mean_infections = mean(num_new_infect),
                             prob_quantiles = quantile(probability_inf, 
                                                       c(0.25, 0.5, 0.75, 
                                                         0.9, 0.95)),
                             inf_quantiles = quantile(num_new_infect, 
                                                      c(0.25, 0.5, 0.75, 
                                                        0.9, 0.95)))))

}

TB <- better_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(TB$prob)
TB_df <- data.frame(prob = TB$prob)

####
Measles <- better_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(Measles$prob)

####
Influenza <- better_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(Influenza$prob)


###
SARS <- better_monte(10000, dis_quanta = 2.7, quanta_sd = 1.2,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(SARS$prob)
```


We had 2 main goals going into this project:

- we wanted to observe probability of infection as we changed some of Beggs et al. set variables such as room volume, number of ppl infected, and exposure time
- we wanted to replicate their Monte Carlo approach and apply it to a different disease (SARS)
- quantum - the amount of airborne material required to infect a majority of susceptible individuals exposed to it


## Scenario 1


For the first scenario, the factors we changed were:

- room size has an increased size
- exposure time is longer by 30 seconds (1 minute in total). 
- only one infected person in the room. 

### Prediction: the infection probability would decrease. 


## Scenario 1 Results

```{r Scenario 1 Graph, echo=FALSE, warning=FALSE}

scenario_monte <- function(m = 10000, dis_quantaS, quanta_sdS, room_vol = 300, 
                         infected_individuals = 1, ACH = 4, t_scen = max(rnorm(1, mean = 1, sd = 0.167))){ 
  prob_inf <- numeric(m)
  num_infec <- numeric(m)
  sus_per <- numeric(m)

  for(i in 1:m) {
    
    p_scen <- max(rnorm(1, mean = 0.48, sd = 0.20), 0)
    phi_scen <- max(rnorm(1, mean = dis_quantaS, sd = quanta_sdS),0)
    S0_scen <- max(round(rnorm(1, mean = 19, sd = 10)),1)
    
    q_2_scen <- infected_individuals * phi_scen
 
    nucci_scen <- ((p_scen*q_2_scen)/room_vol) * 
      ((ACH * t_scen + exp(-ACH * t_scen) -1) / (ACH^2))
    
    S_t_scen <- S0_scen * exp(-nucci_scen)
    
    exp_inf_scen <- S0_scen - S_t_scen
    inf_scen <- rpois(1, lambda = exp_inf_scen)
    
    sus_prob_scen <- inf_scen / S0_scen
    
    prob_inf[i] <- sus_prob_scen
    num_infec[i] <- inf_scen
    sus_per[i] <- S_t_scen
  }
  
  return(list(prob = prob_inf,
              infections = num_infec,
              survivors = sus_per,
              summary = list(mean_prob = mean(prob_inf),
                             mean_infections = mean(num_infec),
                             prob_quantiles = quantile(prob_inf, 
                                                       c(0.25, 0.5, 0.75, 
                                                         0.9, 0.95)),
                             inf_quantiles = quantile(num_infec, 
                                                      c(0.25, 0.5, 0.75, 
                                                        0.9, 0.95)))))

}

scenario_1 <- scenario_monte(10000, dis_quantaS = 12.7, quanta_sdS = 3.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4, t_scen = max(rnorm(1, mean = 1, sd = 0.167)))

# The room volume is larger (300m^3 opposed to 132m^3). The exposure time is longer by 30 minutes (1 hr in total). There is still only one infected person in the room. Before hand, we predicted that the infection probability would decrease. 

scen1_TB <- scenario_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)
scen1_meas <- scenario_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)
scen1_flu <- scenario_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)

TB_df1 <- data.frame(prob=as.numeric(scen1_TB$prob), Disease="TB")
Meas_df1 <- data.frame(prob=as.numeric(scen1_meas$prob), Disease="Measles")
Flu_df1 <- data.frame(prob=as.numeric(scen1_flu$prob), Disease="Influenza")

comb_df1 <- rbind(TB_df1, Meas_df1, Flu_df1)
comb_df1$prob <- pmin(pmax(comb_df1$prob, 0),1)

ggplot(comb_df1, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y=..count.. / sum(..count..) * 100),
                 bins=35, position="identity", alpha=0.4) +
  labs(title="Scenario 1: Probability of Infection Distribution by Disease",
       x="Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  theme_minimal()
```


## Scenario 2


For the second scenario, the factors we changed were:
- room size has a decreased size
- exposure time is set at the same time- 30 seconds 
- now two infected person in the room. 

The room volume was smaller (50m^3 opposed to 132m^2). The exposure time is the same as the article at 30 minutes. There are now two infected people in the room. Before hand, we predicted that the infection probability would increase. 

### Prediction: the infection probability would increase 


## Scenario 2 Results
```{r Scenario 2 Graph, echo=FALSE}
scen2_TB <- scenario_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)
scen2_meas <- scenario_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)
scen2_flu <- scenario_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)

TB_df2 <- data.frame(prob=as.numeric(scen2_TB$prob), Disease="TB")
Meas_df2 <- data.frame(prob=as.numeric(scen2_meas$prob), Disease="Measles")
Flu_df2 <- data.frame(prob=as.numeric(scen2_flu$prob), Disease="Influenza")

comb_df2 <- rbind(TB_df2, Meas_df2, Flu_df2)
comb_df2$prob <- pmin(pmax(comb_df2$prob, 0),1)

ggplot(comb_df2, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y=..count.. / sum(..count..) * 100),
                 bins=35, position="identity", alpha=0.4) +
  labs(title="Scenario 2: Probability of Infection Distribution by Disease",
       x="Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  theme_minimal()
```


## New Disease
The disease we wanted to test everything with was SARS. SARS, or severe acute respiratory syndrome, is another common respiratory illness that is typically spread through the air and being is close contact with an infected person. The quanta was: ; the results were: 

```{r graph, echo=FALSE}

TB_df$Disease <- "TB"
Measles_df <- data.frame(prob=Measles$prob, Disease = "Measles")
Influenza_df <- data.frame(prob=Influenza$prob, Disease = "Influenza")
SARS_df <- data.frame(prob=SARS$prob, Disease="SARS")

combined_df <- rbind(TB_df, Measles_df, Influenza_df, SARS_df)

ggplot(combined_df, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y= ..count.. / sum(..count..) * 100),
                  bins=35, position="identity", alpha=0.4) +
  labs(title="Prob. of Infection Distribution by Disease",
       x = "Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue", "SARS"="yellow3")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue", "SARS"="yellow3")) +
  theme_minimal()

```



## Conclusion

