---
title: "Monte Carlo Applied to Infectious Airborne Diseases"
subtitle: "Stat 400"
author: "Group 1: Robin Thrush and Kaitlynn Walston"
output: 
  ioslides_presentation:
    css: styles.css
---

```{r setup, include=FALSE}
set.seed(445)
knitr::opts_chunk$set(echo = FALSE)

library(dplyr)
library(stats)
library(tidyr)
library(tidyverse)
library(broom)
library(ggplot2)
library(tibble)
```
## Introduction

Stochastic Monte Carlo Model to simulate the probability of a susceptible person
catching an airborne disease.

## Article

Beggs. et al., a study conducted to understand the harm infectious people bring to healthcare waiting rooms. Their
hope was to to have findings that backed up the idea to have more precautions to minimize the spread of airborne infections. The diseases they looked at were Tuberculosis, Measles, and Influenza.

In our article, the researchers attempt to predict the infection rate of an airborne disease when waiting in a waiting room. They did their study on Tuberculous, Measles, and Influenza. They had had specific requirements of room size, ventilation rate, ...

## Assumptions

- Only one infected person in the waiting room
- Once infected, the person can not infect others
- The air being ventilated through the room is mixture of outside air and "infected" air (not being filtered)

## Replication

```{r Replication, echo=FALSE}
better_monte <- function(m = 10000, dis_quanta, quanta_sd, room_vol = 132, 
                         infected_individuals = 1, ACH = 4){ 
  probability_inf <- numeric(m)
  num_new_infect <- numeric(m)
  sus_persons <- numeric(m)

  for(i in 1:m) {
    
    p_2 <- max(rnorm(1, mean = 0.48, sd = 0.20), 0)
    phi_2 <- max(rnorm(1, mean = dis_quanta, sd = quanta_sd),0)
    t_2 <- max(rnorm(1, mean = 0.5, sd = 0.167), 0)
    S0 <- max(round(rnorm(1, mean = 19, sd = 10)),1)
    
    q_2 <- infected_individuals * phi_2
 
    nucci <- ((p_2*q_2)/room_vol) * 
      ((ACH * t_2 + exp(-ACH * 
                                      t_2) -1) / (ACH^2))
    
    S_t <- S0 * exp(-nucci)
    
    exp_inf <- S0 - S_t
    inf <- rpois(1, lambda = exp_inf)
    
    sus_prob <- inf / S0
    
    probability_inf[i] <- sus_prob
    num_new_infect[i] <- inf
    sus_persons[i] <- S_t
  }
  
  return(list(prob = probability_inf,
              infections = num_new_infect,
              survivors = sus_persons,
              summary = list(mean_prob = mean(probability_inf),
                             mean_infections = mean(num_new_infect),
                             prob_quantiles = quantile(probability_inf, 
                                                       c(0.25, 0.5, 0.75, 
                                                         0.9, 0.95)),
                             inf_quantiles = quantile(num_new_infect, 
                                                      c(0.25, 0.5, 0.75, 
                                                        0.9, 0.95)))))

}

TB <- better_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(TB$prob)

TB_df <- data.frame(prob = TB$prob)

TB_plot <- ggplot(data = TB_df, aes(x = prob)) + 
  geom_histogram(aes(y = ..count.. /sum(..count..) *100),
                 bins = 35, 
                 fill = "lightgrey",
                 color = "black") +
  labs(title = "TB Prob. Distrib.",
       x = "Prob. of Infection",
       y = "Percentage of Simulations")


####
Measles <- better_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(Measles$prob)


####
Influenza <- better_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(Influenza$prob)


###
SARS <- better_monte(10000, dis_quanta = 2.7, quanta_sd = 1.2,
                   room_vol = 132, infected_individuals = 1, 
                   ACH = 4)

mean(SARS$prob)
```


We had 2 main going into this project:
- we wanted to test their infection rate they found by changing different variables such as room size and how many people were infected
- we wanted to replicate their Monte Carlo approach and apply it to a different disease (SARS)
- quantum - the amount of airborne material required to infect a majority of susceptible individuals exposed to it

## Scenario 1

The room size is has an increased size than the one in the article. The exposure time is longer by 30 seconds (1 minute in total). There is still only one infected person in the room. Before hand, we predicted that the infection probability would decrease. 

```{r Scenario 1 Graph, echo=FALSE}
scen1_TB <- scenario_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)
scen1_meas <- scenario_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)
scen1_flu <- scenario_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 300, infected_individuals = 1, 
                   ACH = 4)

TB_df1 <- data.frame(prob=as.numeric(scen1_TB$prob), Disease="TB")
Meas_df1 <- data.frame(prob=as.numeric(scen1_meas$prob), Disease="Measles")
Flu_df1 <- data.frame(prob=as.numeric(scen1_flu$prob), Disease="Influenza")

comb_df1 <- rbind(TB_df1, Meas_df1, Flu_df1)
comb_df1$prob <- pmin(pmax(comb_df1$prob, 0),1)

ggplot(comb_df1, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y=..count.. / sum(..count..) * 100),
                 bins=35, position="identity", alpha=0.4) +
  labs(title="Scenario 1: Probability of Infection Distribution by Disease",
       x="Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  theme_minimal()
```


## Scenario 2

The room size was smaller than the one in the article. The exposure time is the same as the article at 30 seconds. There is now two infected people in the room. Before hand, we predicted that the infection probability would increase. 

```{r Scenario 2 Graph, echo=FALSE}
scen2_TB <- scenario_monte(10000, dis_quanta = 12.7, quanta_sd = 3.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)
scen2_meas <- scenario_monte(10000, dis_quanta = 100, quanta_sd = 25.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)
scen2_flu <- scenario_monte(10000, dis_quanta = 570, quanta_sd = 173.0,
                   room_vol = 50, infected_individuals = 2, 
                   ACH = 4)

TB_df2 <- data.frame(prob=as.numeric(scen2_TB$prob), Disease="TB")
Meas_df2 <- data.frame(prob=as.numeric(scen2_meas$prob), Disease="Measles")
Flu_df2 <- data.frame(prob=as.numeric(scen2_flu$prob), Disease="Influenza")

comb_df2 <- rbind(TB_df2, Meas_df2, Flu_df2)
comb_df2$prob <- pmin(pmax(comb_df2$prob, 0),1)

ggplot(comb_df2, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y=..count.. / sum(..count..) * 100),
                 bins=35, position="identity", alpha=0.4) +
  labs(title="Scenario 2: Probability of Infection Distribution by Disease",
       x="Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue")) +
  theme_minimal()
```





## New Disease
The disease we wanted to test everything with was SARS. SARS, or severe acute respiratory syndrome, is another common respiratory illness that is typically spread through the air and being is close contact with an infected person. The quanta was: ; the results were: 

```{r graph, echo=FALSE}

TB_df$Disease <- "TB"
Measles_df <- data.frame(prob=Measles$prob, Disease = "Measles")
Influenza_df <- data.frame(prob=Influenza$prob, Disease = "Influenza")
SARS_df <- data.frame(prob=SARS$prob, Disease="SARS")

combined_df <- rbind(TB_df, Measles_df, Influenza_df, SARS_df)

ggplot(combined_df, aes(x=prob, fill=Disease, color=Disease)) +
  geom_histogram(aes(y= ..count.. / sum(..count..) * 100),
                  bins=35, position="identity", alpha=0.4) +
  labs(title="Prob. of Infection Distribution by Disease",
       x = "Probability of Infection",
       y = "Percentage of Simulations") +
  scale_color_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue", "SARS"="yellow3")) +
  scale_fill_manual(values = c("TB" = "green", "Measles" = "magenta",
                               "Influenza"="blue", "SARS"="yellow3")) +
  theme_minimal()

```



## Conclusion

